---
title: "Reproducible Research Project 1"
author: "Adraine Upshaw"
date: "December 11, 2015"
output: html_document
fontsize: 10pt
---
Loading and preprocessing the data
-------------------------------------
```{r}
setwd("~/DataScience/Reproducible Research")
data <- read.csv("activity.csv")
head(data)
summary(data)
str(data)
data$date2 <- as.Date(data$date,format="%Y-%m-%d")
class(data$date2)
```

<<<<<<< HEAD
Loading and preprocessing the data
-------------------------------------
```{r}
options(scipen = 2, digits = 0)
setwd("~/DataScience/Reproducible Research")
data <- read.csv("activity.csv")
head(data)
summary(data)
str(data)
data$date2 <- as.Date(data$date,format="%Y-%m-%d")
class(data$date2)
```

What is mean total number of steps taken per day?
-------------------------------------
1. Calculate the total number of steps taken per day
```{r,warning=FALSE, message=FALSE,}
library(plyr);library(dplyr)

data_steps <- data[,c("date2","steps")]
group_steps <- group_by(data_steps,date2)
total_steps <- summarise_each(group_steps,funs(n(),sum))
head(total_steps)
```

2. Make a histogram of the total number of steps taken each day
```{r,warning=FALSE, message=FALSE, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
myhist <-hist(total_steps$sum,xlab="Total Steps",main="Histogram of Total Steps")
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r computemean}
library(knitr);
total_mean <- round(mean(total_steps$sum,na.rm =TRUE ),digits = 2)
total_med <- median(total_steps$sum,na.rm =TRUE )
```
The mean of the total number of steps taken per day is `r total_mean`.
The median of the total number of steps taken per day is `r total_med`.

What is the average daily activity pattern?
-------------------------------------
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r,warning=FALSE, message=FALSE,, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
total_interval <- aggregate(steps ~ interval, data, mean)
mymean <- mean(total_interval$steps)
plot(total_interval$interval,total_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
abline(h=mymean,col = "red", lwd = 1)
leg.txt <- c("mean steps")
legend("topright",      # Set location of the legend
       cex=0.75,     #everything is small
       legend = leg.txt,               # Specify text 
       col = c("red"),     # Set colors for legend
       lty = c(2),                   # Set type of lines in legend
       merge = TRUE)
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r,warning=FALSE, message=FALSE,}
max_interval <- arrange(total_interval,desc(steps))
max <- max_interval[1,]
=======
What is mean total number of steps taken per day?
-------------------------------------
1. Calculate the total number of steps taken per day
```{r,warning=FALSE, message=FALSE,}
library(plyr);library(dplyr)

data_steps <- data[,c("date2","steps")]
group_steps <- group_by(data_steps,date2)
total_steps <- summarise_each(group_steps,funs(n(),sum))
head(total_steps)
```

2. Make a histogram of the total number of steps taken each day
```{r,warning=FALSE, message=FALSE, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
myhist <-hist(total_steps$sum,xlab="Total Steps",main="Histogram of Total Steps")
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r computemean}
library(knitr);options(digits=2)
total_mean <- round(mean(total_steps$sum,na.rm =TRUE ),digits = 2)
total_med <- median(total_steps$sum,na.rm =TRUE )
```
The mean of the total number of steps taken per day is `r total_mean`.
The median of the total number of steps taken per day is `r total_med`.

What is the average daily activity pattern?
-------------------------------------
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r,warning=FALSE, message=FALSE,, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
total_interval <- aggregate(steps ~ interval, data, mean)
mymean <- mean(total_interval$steps)
plot(total_interval$interval,total_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
abline(h=mymean,col = "red", lwd = 1)
leg.txt <- c("mean steps")
legend("topright",      # Set location of the legend
       cex=0.75,     #everything is small
       legend = leg.txt,               # Specify text 
       col = c("red"),     # Set colors for legend
       lty = c(2),                   # Set type of lines in legend
       merge = TRUE)
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r,warning=FALSE, message=FALSE,}
max_interval <- arrange(total_interval,desc(steps))
max <- max_interval[1,]

```
The 5-minute interval containing the maximum number of steps is `r max[1,1]` with `r round(max[1,2],2)` average steps.
>>>>>>> 087e262aad7758c737895d4c38484abc1a6aaff3

```
The 5-minute interval containing the maximum number of steps is `r max[1,1]` with `r round(max[1,2],2)` average steps.

<<<<<<< HEAD

Imputing missing values
-------------------------------------
1. Calculate and report the total number of missing values in the dataset
```{r,warning=FALSE, message=FALSE}
missing <- sum(!complete.cases(data))
missing
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
```{r,warning=FALSE, message=FALSE}

library(zoo)
data$steps2 <- na.aggregate(data$steps)
missing_check <- sum(is.na(data$steps2))
missing_check
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r,warning=FALSE, message=FALSE}
data2 <- select(data,date2,interval,steps2)

```

4.1. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r,warning=FALSE, message=FALSE,, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
total <- aggregate(steps2 ~ date2, data2, sum)
myhist2 <-hist(total$steps2,xlab="Total Steps",main="Histogram of Total Steps after Replacing Missing Values")
```

=======
Imputing missing values
-------------------------------------
1. Calculate and report the total number of missing values in the dataset
```{r,warning=FALSE, message=FALSE}
missing <- sum(!complete.cases(data))
missing
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
```{r,warning=FALSE, message=FALSE}

library(zoo)
data$steps2 <- na.aggregate(data$steps)
missing_check <- sum(is.na(data$steps2))
missing_check
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r,warning=FALSE, message=FALSE}
data2 <- select(data,date2,interval,steps2)

```

4.1. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r,warning=FALSE, message=FALSE,, fig.width=5, fig.height=4,fig.align='center',strip.white=FALSE}
total <- aggregate(steps2 ~ date2, data2, sum)
myhist2 <-hist(total$steps2,xlab="Total Steps",main="Histogram of Total Steps after Replacing Missing Values")
```

>>>>>>> 087e262aad7758c737895d4c38484abc1a6aaff3
4.2. Do these values differ from the estimates from the first part of the assignment? 
```{r computemean2}
total_mean2 <- round(mean(total$steps2,na.rm =TRUE ),digits = 2)
total_med2 <- median(total$steps2,na.rm =TRUE )
```
The mean of the total number of steps taken per day after imputing is `r round(total_mean2)`.
The median of the total number of steps taken per day after imputing is `r round(total_med2)`.

4.3. What is the impact of imputing missing data on the estimates of the total daily number of steps?
By imputing the average, the distrubution is now not skewed (mean=median)

Are there differences in activity patterns between weekdays and weekends?
-------------------------------------
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r,warning=FALSE, message=FALSE}
data3 <- mutate(data2,day_type=ifelse(weekdays(data2$date2) == "Saturday" | weekdays(data2$date2) == "Sunday", "weekend", "weekday"))
table(data3$day_type)

```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r,warning=FALSE, message=FALSE,, fig.width=6, fig.height=6,fig.align='center',strip.white=FALSE}
library(ggplot2)
plot_data <- aggregate(steps2~interval+day_type, data =data3, mean)

ggplot(plot_data, aes(x=interval, y=steps2, color = day_type)) +
  geom_line() +
  facet_wrap(~day_type, ncol = 1, nrow=2)

```